### 2019-11-04 submission/manuscript.Rmd
Response to reviewers

Few scientists set out on a nearly two decade-long journey with a specific goal in mind. Often we fail to start a scientific journey because it looks too hard. Perhaps we get discouraged by all of the things that could go wrong. Maybe we stray from the path because we find something that is more interesting. Every scientist picks their own path and takes their own forks in the road. From the outside, it may appear to be a random walk. Nevertheless, these meanderings are common in science.

@@

With the growth in popularity of 16S rRNA gene sequencing there has naturally been an expansion in the number of people developing tools to analyze these data. Months after the paper describing mothur was published, the paper describing QIIME was published [@Caporaso2010a]. Over the past 10 years, many have attempted to create analogies comparing the two programs: Pepsi vs Coke, Apple vs Windows, etc. It is never clear which software is which brand and whether the comparisons are meant as a complement or an insult. Regardless, both programs are very popular. From my perspective, most of the differences are cosmetic (http://blog.mothur.org/2016/01/12/mothur-and-qiime/). To me the most meaningful difference between mothur and QIIME is the choice of algorithms used to cluster sequences into OTUs. QIIME's advocacy for open and closed-reference clustering and USEARCH or VSEARCH-based *de novo* clustering results in lower quality OTU assignments relative to the *de novo* clustering algorithms available within mothur [@Westcott2015; @Westcott2017]. QIIME is set of wrapper scripts that help users to transition data between independent packages. For example, with QIIME (through version 1.9.1), it was even possible to run mothur through QIIME. One can also run the \text{na\"ive} Bayesian classifier through QIIME using the original code developed by the RDP. This caused great frustration for many users because there were numerous software dependencies that had to be installed. Although the QIIME developers would go on to create virtual machines and use packaging tools to simplify installation, these fixes required sophistication by users who we knew struggled with the basics of navigating a command line. In contrast, when a user runs mothur, they are running mothur. The \text{na\"ive} Bayesian classifier code that is in mothur is a rewritten version of the original code. When we rewrite someone's software we do it with an eye to improving performance, access, and utility for non-16S rRNA gene sequence data. For example, while 454 data was popular, PyroNoise was an effective tool for denoising flowgram data [@Quince2009]. Running the original code required a large Linux computer cluster and knowledge of bash and Perl scripting. When we rewrote the code for mothur, we made it accessible to people using any operating system with a simple command interface (i.e. trim.flows and shhh.flows). Our approach requires significant developer effort, but saves considerable user effort. As this benefit is multiplied across thousands of projects, the savings to users has been considerable.

>>>>

@@

With the growth in popularity of 16S rRNA gene sequencing there has naturally been an expansion in the number of people developing tools to analyze these data. Months after the paper describing mothur was published, the paper describing QIIME was published [@Caporaso2010a]. Over the past 10 years, many have attempted to create analogies comparing the two programs: Pepsi vs Coke, Apple vs Windows, etc. It is never clear which software is which brand and whether the comparisons are meant as a complement or an insult. Regardless, both programs are very popular. From my perspective, most of the differences are cosmetic (http://blog.mothur.org/2016/01/12/mothur-and-qiime/). To me the most meaningful difference between mothur and QIIME is the choice of algorithms used to cluster sequences into OTUs. QIIME's advocacy for open and closed-reference clustering and USEARCH or VSEARCH-based *de novo* clustering results in lower quality OTU assignments relative to the *de novo* clustering algorithms available within mothur [@Westcott2015; @Westcott2017]. QIIME is set of wrapper scripts that help users to transition data between independent packages. For example, with QIIME (through version 1.9.1), it was even possible to run mothur through QIIME. One can also run the \text{na\"ive} Bayesian classifier through QIIME using the original code developed by the RDP. Structuring QIIME as a set of wrappers caused great frustration for many users because there were numerous software dependencies that had to be installed. The benefits included the ability for users to access to a wider set of tools and for developers to tie their tool into the popular software package. Although the QIIME developers would go on to create virtual machines and use packaging tools to simplify installation, these fixes required sophistication by users who we knew struggled with the basics of navigating a command line. In contrast, when a user runs mothur, they are running mothur. The \text{na\"ive} Bayesian classifier code that is in mothur is a rewritten version of the original code. When we rewrite someone's software we do it with an eye to improving performance, access, and utility for non-16S rRNA gene sequence data. For example, while 454 data was popular, PyroNoise was an effective tool for denoising flowgram data [@Quince2009]. Running the original code required a large Linux computer cluster and knowledge of bash and Perl scripting. When we rewrote the code for mothur, we made it accessible to people using any operating system with a simple command interface (i.e. trim.flows and shhh.flows). Our approach requires significant developer effort, but saves considerable user effort. As this benefit is multiplied across thousands of projects, the savings to users has been considerable.

### 2019-11-04 submission/manuscript.tex
Response to reviewers

With the growth in popularity of 16S rRNA gene sequencing there has
naturally been an expansion in the number of people developing tools to
analyze these data. Months after the paper describing mothur was
published, the paper describing QIIME was published (46). Over the past
10 years, many have attempted to create analogies comparing the two
programs: Pepsi vs Coke, Apple vs Windows, etc. It is never clear which
software is which brand and whether the comparisons are meant as a
complement or an insult. Regardless, both programs are very popular.
From my perspective, most of the differences are cosmetic
(\url{http://blog.mothur.org/2016/01/12/mothur-and-qiime/}). To me the
most meaningful difference between mothur and QIIME is the choice of
algorithms used to cluster sequences into OTUs. QIIME's advocacy for
open and closed-reference clustering and USEARCH or VSEARCH-based
\emph{de novo} clustering results in lower quality OTU assignments
relative to the \emph{de novo} clustering algorithms available within
mothur (47, 48). QIIME is set of wrapper scripts that help users to
transition data between independent packages. For example, with QIIME
(through version 1.9.1), it was even possible to run mothur through
QIIME. One can also run the \text{na\"ive} Bayesian classifier through
QIIME using the original code developed by the RDP. This caused great
frustration for many users because there were numerous software
dependencies that had to be installed. Although the QIIME developers
would go on to create virtual machines and use packaging tools to
simplify installation, these fixes required sophistication by users who
we knew struggled with the basics of navigating a command line. In
contrast, when a user runs mothur, they are running mothur. The
\text{na\"ive} Bayesian classifier code that is in mothur is a rewritten
version of the original code. When we rewrite someone's software we do
it with an eye to improving performance, access, and utility for non-16S
rRNA gene sequence data. For example, while 454 data was popular,
PyroNoise was an effective tool for denoising flowgram data (49).
Running the original code required a large Linux computer cluster and
knowledge of bash and Perl scripting. When we rewrote the code for
mothur, we made it accessible to people using any operating system with
a simple command interface (i.e.~trim.flows and shhh.flows). Our
approach requires significant developer effort, but saves considerable
user effort. As this benefit is multiplied across thousands of projects,
the savings to users has been considerable.

>>>>

With the growth in popularity of 16S rRNA gene sequencing there has
naturally been an expansion in the number of people developing tools to
analyze these data. Months after the paper describing mothur was
published, the paper describing QIIME was published (46). Over the past
10 years, many have attempted to create analogies comparing the two
programs: Pepsi vs Coke, Apple vs Windows, etc. It is never clear which
software is which brand and whether the comparisons are meant as a
complement or an insult. Regardless, both programs are very popular.
From my perspective, most of the differences are cosmetic
(\url{http://blog.mothur.org/2016/01/12/mothur-and-qiime/}). To me the
most meaningful difference between mothur and QIIME is the choice of
algorithms used to cluster sequences into OTUs. QIIME's advocacy for
open and closed-reference clustering and USEARCH or VSEARCH-based
\emph{de novo} clustering results in lower quality OTU assignments
relative to the \emph{de novo} clustering algorithms available within
mothur (47, 48). QIIME is set of wrapper scripts that help users to
transition data between independent packages. For example, with QIIME
(through version 1.9.1), it was even possible to run mothur through
QIIME. One can also run the \text{na\"ive} Bayesian classifier through
QIIME using the original code developed by the RDP. Structuring QIIME as
a set of wrappers caused great frustration for many users because there
were numerous software dependencies that had to be installed. The
benefits included the ability for users to access to a wider set of
tools and for developers to tie their tool into the popular software
package. Although the QIIME developers would go on to create virtual
machines and use packaging tools to simplify installation, these fixes
required sophistication by users who we knew struggled with the basics
of navigating a command line. In contrast, when a user runs mothur, they
are running mothur. The \text{na\"ive} Bayesian classifier code that is
in mothur is a rewritten version of the original code. When we rewrite
someone's software we do it with an eye to improving performance,
access, and utility for non-16S rRNA gene sequence data. For example,
while 454 data was popular, PyroNoise was an effective tool for
denoising flowgram data (49). Running the original code required a large
Linux computer cluster and knowledge of bash and Perl scripting. When we
rewrote the code for mothur, we made it accessible to people using any
operating system with a simple command interface (i.e.~trim.flows and
shhh.flows). Our approach requires significant developer effort, but
saves considerable user effort. As this benefit is multiplied across
thousands of projects, the savings to users has been considerable.
